---
title: "UN Votes"
author: ""
date: "08.17.20"
output: 
  html_document: 
    toc: true
    toc_float: true
---

We want to look at the voting history of countries in the United Nations General 
Assembly using data from package `unvotes`. 

```{r set-up, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      echo = FALSE)
```

```{r load-packages, warning=FALSE, message=FALSE}
library(unvotes)
library(tidyverse)
library(lubridate)
library(DT)
```

## Data

We will work with three data sets: `un_roll_calls`, `un_roll_call_issues`, 
and `un_votes`. Each data set contains a variable called `rcid`, the roll call 
id, which can be used to join the data sets with one another.

- The `un_votes` data set provides information on the voting history of the 
  United Nations General Assembly. It contains one row for each country/vote 
  pair.

```{r view-un_votes}
un_votes
```

- The `un_roll_calls` data set contains information on each roll call vote of 
  the United Nations General Assembly.

```{r view-un_roll_calls}
un_roll_calls
```

- The `un_roll_call_issues` data set contains issue classifications of roll 
  call votes of the United Nations General Assembly. There are many votes that 
  have no issue classification, and some are classified under more than 
  one issue.

```{r view-un_roll_call_issues}
un_roll_call_issues
```

## Analysis

### Part 1

We begin by looking at how often each country voted "yes" on a resolution in 
each year. We'll visualize the results, so let's pick a few countries of 
interest and focus our analysis on them.

```{r country-list}
country_list <- c("United States of America", "Ghana", "Mexico", "France")
```

```{r plot-yearly-yes, fig.width=10, warning=FALSE}
un_votes %>%
  filter(country %in% country_list) %>%
  inner_join(un_roll_calls, by = "rcid") %>% 
  group_by(year = year(date), country) %>%
  summarize(votes = n(),
            percent_yes = mean(vote == "yes")) %>%
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_line() +
    ylab("% of votes that are 'Yes'")
```

### Part 2

Next, let's see how the voting records have changed over the years on each of the issues.

```{r plot-yearly-yes-issue, fig.width=10, message = FALSE, warning=FALSE}
un_votes %>%
  filter(country %in% c("United States of America", "France")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  group_by(country, year = year(date), issue) %>%
  summarize(votes = n(),
            percent_yes = mean(vote == "yes")) %>%
  filter(votes > 5) %>%  # Only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) + 
    ylab("% of votes that are 'Yes'") +
    facet_wrap(~ issue)
```

### Discussion Questions

1. Consider the plot from Part 1. Describe how the voting behaviors of the four  countries have changed over time.

2. Consider the plot from Part 2. 
    - On which issues have the two countries voted most similarly in 
      recent years?  
    - On which issues have they voted most differently in recent years? 
    - Has this changed over time?


## References

1. David Robinson (2017). unvotes: United Nations General Assembly Voting Data. 
   R package version 0.2.0. https://CRAN.R-project.org/package=unvotes.
2. Erik Voeten "Data and Analyses of Voting in the UN General Assembly" 
   Routledge Handbook of International Organization, edited by Bob Reinalda 
   (published May 27, 2013).
3. This assignment was adapted from 
   [UN Votes](https://github.com/rstudio-education/datascience-box/tree/master/appex/ae-01-un-votes) 
   exercise and the examples presented in the 
   [unvotes package vignette](https://cran.r-project.org/web/packages/unvotes/vignettes/unvotes.html).


## Appendix

Below is a list of countries in the data set:

```{r}
un_votes %>% 
  arrange(country) %>% 
  select(country) %>%
  distinct() %>%
  datatable()
```
