---
title: "The Central Limit Theorem"
subtitle: "(CLT)"
author: "Prof. Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
---



layout: true

<div class="my-footer">
<span>
<a href="http://datasciencebox.org" target="_blank">datasciencebox.org</a>
</span>
</div> 

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 3, fig.width = 5, dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center") 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(emo)
library(openintro)
library(infer)
library(patchwork)
```

---

class: middle center

## [Click for PDF of slides](13-clt.pdf)

---


class: center, middle

## Sample Statistics and Sampling Distributions

---

## Variability of sample statistics

- We've seen that each sample from the population yields a slightly different 
sample statistic (sample mean, sample proportion, etc.)

- Previously we've quantified this value via simulation

- Today we talk about some of the theory underlying .vocab[sampling distributions],
particularly as they relate to sample means.

---

## Statistical inference

- Statistical inference is the act of generalizing from a sample in order to 
make conclusions regarding a population. 

- We are interested in population parameters, which we do not observe. Instead, 
we must calculate statistics from our sample in order to learn about them.

- As part of this process, we must quantify the degree of uncertainty in our 
sample statistic. 

---

## Sampling distribution of the mean

Suppose we’re interested in the mean resting heart rate of students at Duke, and 
are able to do the following:

--

1. Take a random sample of size $n$ from this population, and calculate the 
mean resting heart rate in this sample, $\bar{X}_1$

--

2. Put the sample back, take a second random sample of size $n$, and calculate 
the mean resting heart rate from this new sample, $\bar{X}_2$

--

3. Put the sample back, take a third random sample of size $n$, and calculate
the mean resting heart rate from this sample, too...

--

...and so on.

---

## Sampling distribution of the mean

After repeating this many times, we have a dataset that has the
sample averages from the population: $\bar{X}_1$, $\bar{X}_2$, $\cdots$,
$\bar{X}_K$ (assuming we took $K$ total samples).

--

.question[
Can we say anything about the distribution of these sample means (that is, the
.vocab[sampling distribution] of the mean?) 
]

*(Keep in mind, we don't know what the underlying distribution of mean resting 
heart rate looks like in Duke students!)*

---

class: center, middle

## The Central Limit Theorem

---

class: middle

A quick caveat...

For now, let's assume we know the underlying standard deviation, $\sigma$, from our distribution 

---

## The Central Limit Theorem

For a population with a well-defined mean $\mu$ and standard deviation $\sigma$,
these three properties hold for the distribution of sample average $\bar{X}$,
assuming certain conditions hold:

--

1. The mean of the sampling distribution of the mean is identical to the 
population mean $\mu$.

--

2. The standard deviation of the distribution of the sample averages is
$\sigma/\sqrt{n}$.
  - This is called the .vocab[standard error] (SE) of the mean. 
 
--

3. For $n$ large enough, the shape of the
sampling distribution of means is approximately .vocab[normally distributed].

---

## The normal (Gaussian) distribution

The normal distribution is unimodal and symmetric and is described by its
.vocab[density function]:

If a random variable $X$ follows the normal distribution, then
$$f(x) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left\{ -\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2} \right\}$$
where $\mu$ is the mean and $\sigma^2$ is the variance $(\sigma \text{ is the standard deviation})$

.alert[
We often write $N(\mu, \sigma)$ to describe this distribution.
]
---

## The normal distribution (graphically)

```{r echo = F, fig.height = 2.5}
ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(-3, 3), color = "black", size = 1.5) +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

---

## Wait, *any* distribution?

The central limit theorem tells us that *<b>sample averages</b>* are normally distributed, if we have enough data and certain assumptions hold. 

This is true *even if our original variables are not normally distributed*.

Click [here](http://onlinestatbook.com/stat_sim/sampling_dist/index.html) to see an interactive demonstration of this idea.

---

## Conditions for CLT 

We need to check two conditions for CLT to hold: independence, sample size/distribution. 

--

`r emo::ji("white_check_mark")` .vocab[Independence:] The sampled observations must be independent. This is 
difficult to check, but the following are useful guidelines:

- the sample must be randomly taken
- if sampling without replacement, sample size must be less than 10% of the 
    population size

--

If samples are independent, then by definition one sample's value does not "influence" another sample's value.

---

## Conditions for CLT 

`r emo::ji("white_check_mark")` .vocab[Sample size / distribution:] 

- if data are numerical, usually n > 30 is considered a large enough sample for the CLT to kick in
- if we know for sure that the underlying data are normally distributed, then the distribution of sample averages will also be exactly normal, regardless of the sample size
- if data are categorical, at least 10 successes and 10 failures.

---

class: middle, center

## Let's run our own simulation

---

### Underlying population (not observed in real life!)

.small[
```{r}
rs_pop <- tibble(x = rbeta(100000, 1, 5) * 100)
```
]

```{r echo=FALSE, fig.height=2}
ggplot(data = rs_pop, aes(x = x)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "darkblue") +
  labs(title = "Population distribution", x = "") +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

**The true population parameters**
.small[
```{r echo = F}
rs_pop %>%
  summarise(mu = mean(x), sigma = sd(x))
```
]

---

## Sampling from the population - 1

```{r}
set.seed(1)
samp_1 <- rs_pop %>%
  sample_n(size = 50) %>%
  summarise(x_bar = mean(x))
```

```{r}
samp_1
```

---

## Sampling from the population - 2

```{r}
set.seed(2)
samp_2 <- rs_pop %>%
  sample_n(size = 50) %>% 
  summarise(x_bar = mean(x))
```

```{r}
samp_2
```
---

## Sampling from the population - 3

```{r}
set.seed(3)
samp_3 <- rs_pop %>%
  sample_n(size = 50) %>% 
  summarise(x_bar = mean(x))
```

```{r}
samp_3
```

--

keep repeating...

---

## Sampling distribution

.small[
```{r}
set.seed(092620)
sampling <- rs_pop %>%
  rep_sample_n(size = 50, replace = TRUE, reps = 5000) %>%
  group_by(replicate) %>%
  summarise(xbar = mean(x))
```
]

```{r echo=FALSE, fig.height=1.9}
ggplot(data = sampling, aes(x = xbar)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "darkblue") +
  labs(title = "Sampling distribution of sample means") +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())
```

```{r}
sampling %>%
  summarise(mean = mean(xbar), se = sd(xbar))
```

---

.question[
How do the shapes, centers, and spreads of these distributions compare?
]

```{r echo=FALSE, fig.height=2.5}
p1 <- ggplot(data = rs_pop, aes(x = x)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "darkblue") +
  labs(title = "Population distribution", x = "") +
  xlim(-5, 100) +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())

p2 <- ggplot(data = sampling, aes(x = xbar)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "darkblue") +
  labs(title = "Sampling distribution of sample means", x = "Sample means") +
  xlim(-5, 100) +
  labs(x = " ", 
       y = " ") +
    theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())

p1 / p2
```

---

## Recap

- If certain assumptions are satisfied, regardless of the shape of the 
population distribution, the sampling distribution of the mean follows an 
approximately normal distribution.

--

- The center of the sampling distribution is at the center of the population 
distribution.

--

- The sampling distribution is less variable than the population distribution 
(and we can quantify by how much).

--

.question[
What is the standard error, and how are the standard error and sample size 
related? What does that say about how the spread of the sampling distribution
changes as $n$ increases?
]

---

class: center, middle

## Finding probabilities in R

---

## Probabilities under N(0,1) curve

```{r}
# P(Z < -1.5)
pnorm(-1.5)
```

```{r echo = F, fig.height = 2}
ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#00998a", xlim = c(-3, -1.5), color = "black") +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(-1.5, 3), color = "black") +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

---

## Probability between two values

.question[
If $Z \sim N(0, 1)$, what is $P(-1 < Z < 2)$?
]

--

```{r echo = F, fig.height = 2}
from.z <- qnorm(pnorm(-1))
to.z <- qnorm(pnorm(2))

ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#00998a", xlim = c(from.z, to.z), color = "black") +
    geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(-3, from.z), color = "black") +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(to.z, 3), color = "black") +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

---

##  Probability between two values

.question[
If $Z \sim N(0, 1)$, what is $P(-1 < Z < 2)$?
]


```{r echo = F, fig.height = 2}
from.z <- -3
to.z <- qnorm(pnorm(2))

ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#C71585", xlim = c(from.z, to.z), color = "black") +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(to.z, 3), color = "black") +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

---

## Probability between two values

.question[
If $Z \sim N(0, 1)$, what is $P(-1 < Z < 2)$?
]

```{r echo = F, fig.height = 2}
from.z <- -3
to.z <- qnorm(pnorm(-1))

ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#C71585", xlim = c(from.z, to.z), color = "black") +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(to.z, 3), color = "black") +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

---

## Probability between two values

.question[
If $Z \sim N(0, 1)$, what is $P(-1 < Z < 2)$?
]

```{r echo = F, fig.height = 2}
from.z <- qnorm(pnorm(-1))
to.z <- qnorm(pnorm(2))

ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#00998a", xlim = c(from.z, to.z), color = "black") +
    geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(-3, from.z), color = "black") +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(to.z, 3), color = "black") +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

```{r}
pnorm(2) - pnorm(-1)
```

---

##  Probability between two values

.question[
If $Z \sim N(0, 1)$, what is $P(-1 < Z < 2)$?
]

```{r}
pnorm(2) - pnorm(-1)
```

---

## Finding cutoff values under N(0,1) curve

```{r}
# find Q1
qnorm(0.25)
```


```{r echo = F, fig.height = 2}
from.z <- -3
to.z <- qnorm(.25)
 
ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, fill = "#00998a", xlim = c(from.z, to.z), color = "black") +
  geom_area(stat = "function", fun = dnorm, fill = "#FFFFFF", xlim = c(to.z, 3), color = "black") +
  labs(x = "", 
       y = "Density") +
  theme_bw() +
  scale_x_continuous(breaks = c(-3,-2,-1,0,1,2,3))
```

---

## Looking ahead...

We will use the Central Limit Theorem and the normal distribution to conduct statistical inference.