---
title: "Text analysis"
author: "Prof. Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
---


layout: true

<div class="my-footer">
<span>
<a href="http://datasciencebox.org" target="_blank">datasciencebox.org</a>
</span>
</div> 

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 3, fig.width = 5, dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center") 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(emo)
library(openintro)
library(infer)
library(patchwork)
library(kableExtra)
```

---

class: middle center

## [Click for PDF of slides](22-text-analysis.pdf)

---


## Packages

In addition to `tidyverse` we will be using a few other packages today

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(genius) # https://github.com/JosiahParry/genius
```

---



## Tidy Data

.question[
What makes a data frame tidy?
]

--

1. Each variable must have its own column.
2. Each observation must have its own row.
3. Each value must have its own cell.

---


## Tidytext

- Using tidy data principles can make many text mining tasks easier, more 
effective, and consistent with tools already in wide use.

- Learn more at https://www.tidytextmining.com/.

---

## What is tidy text?

```{r, message=FALSE, warning=FALSE}
text <- c("On your mark ready set let's go", 
          "dance floor pro",
          "I know you know I go psycho", 
          "When my new joint hit", 
          "just can't sit",
          "Got to get jiggy wit it", 
          "ooh, that's it")

text
```

---

## What is tidy text?

```{r, message=FALSE, warning=FALSE}
text_df <- tibble(line = 1:7, text = text)

text_df
```

---

## What is tidy text?

```{r, message=FALSE, warning=FALSE}
text_df %>%
  unnest_tokens(word, text)
```

---

## Let's get some data

We'll use the `genius` package to get song lyric data from [Genius](https://genius.com/).

- `genius_album()` allows you to download the lyrics for an entire album in a 
tidy format. 

- Input: Two arguments: artist and album. Supply the quoted name of artist 
and the album (if it gives you issues check that you have the album name and 
artists as specified on [Genius](https://genius.com/)).

- Output: A tidy data frame with three columns corresponding to the track name, 
the track number, and lyrics

---

## Let's get some data

```{r, message = F, warning = F, cache=TRUE}
tswift <- genius_album(
  artist = "Taylor Swift", 
  album = "Lover"
  )

tswift
```

---

## What songs are in the album?

```{r, message = F, warning = F}
tswift %>%
  distinct(track_title)
```

---

## How long are the songs? 

Length is measured by number of lines

```{r, message = F, warning = F}
tswift %>%
  count(track_title, sort = TRUE)
```

---

## Tidy up your lyrics!

```{r, message = F, warning = F}
tswift_lyrics <- tswift %>%
  unnest_tokens(word, lyric)

tswift_lyrics
```

---

## What are the most common words?

```{r, message = F, warning = F}
tswift_lyrics %>%
  count(word) %>%
  arrange(desc(n))
```

---

## Stop words

- In computing, stop words are words which are filtered out before or after 
processing of natural language data (text).

- They usually refer to the most common words in a language, but there is not a 
single list of stop words used by all natural language processing tools.


---

## What are the most common words?

```{r, message = F, warning = F}
tswift_lyrics %>%
  anti_join(stop_words) %>%
  count(word) %>%
  arrange(desc(n))
```



---


## What are the most common words?

```{r, echo=FALSE, message=FALSE, fig.height = 5.75, fig.width = 11}
tswift_lyrics %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(word) %>%
  arrange(desc(n)) %>%
  top_n(20) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
    geom_col() +
    coord_flip() + 
    theme_minimal() +
    labs(title = "Frequency of 'Lover' lyrics",
         y = "",
         x = "")
```


---

## ...the code

```{r, eval  = F}
tswift_lyrics %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(word) %>%
  arrange(desc(n)) %>%
  top_n(20) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
    geom_col() +
    coord_flip() + 
    theme_minimal() +
    labs(title = "Frequency of 'Lover' lyrics",
         y = "",
         x = "")
```

---

class: middle, center

## Sentiment analysis

---

## Sentiment analysis

- One way to analyze the sentiment of a text is to consider the text as a 
combination of its individual words 

- The sentiment content of the whole text as the sum of the sentiment content of
the individual words

- The sentiment attached to each word is given by a *lexicon*, which may be 
downloaded from external sources

---

## Sentiment lexicons

.pull-left[
```{r}
get_sentiments("afinn")
```
]

--

.pull-right[
```{r}
get_sentiments("bing") 
```
]

---

## Sentiment lexicons

.pull-left[
```{r}
get_sentiments("nrc")
```
]

--

.pull-right[
```{r}
get_sentiments("loughran") 
```
]

---

## Notes about sentiment lexicons

- Not every word is in a lexicon!

```{r}
get_sentiments("bing") %>% 
  filter(word == "data")
```

--

- Lexicons do not account for qualifiers before a word (e.g., "not happy") because they were constructed for one-word tokens only

--

- Summing up each word's sentiment may result in a neutral sentiment, even if there are strong positive and negative sentiments in the body

---

## Sentiments in lyrics

```{r, message = F, warning = F}
tswift_lyrics %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, word, sort = TRUE) 
```

---

## Let's visualize T.Swift's top 10 sentiments

```{r}
tswift_top10 <- tswift_lyrics %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, word) %>%
  arrange(desc(n)) %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup()
```

---

## Visualizing the top 10

.midi[
```{r, echo=FALSE, message=FALSE, fig.height = 5.75, fig.width = 11}
ggplot(tswift_top10, aes(fct_reorder(word, n), n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ sentiment, scales = "free") +
  theme_minimal() +
  labs(title = "Sentiments in Taylor Swift Lyrics",
         x = "", y = "")
```
]

---

## Let's remove the redundant legend

```{r, message=FALSE, fig.height = 4, eval = F}
ggplot(tswift_top10, aes(fct_reorder(word, n), n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ sentiment, scales = "free") +
  theme_minimal() +
  labs(title = "Sentiments in Taylor Swift Lyrics", x = "", y = "") + 
  guides(fill = FALSE) #<<
```

---

## Let's remove the redundant legend

```{r, echo = F, message=FALSE, fig.height = 2.5}
ggplot(tswift_top10, aes(fct_reorder(word, n), n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ sentiment, scales = "free") +
  theme_minimal() +
  labs(title = "Sentiments in Taylor Swift Lyrics", x = "", y = "") + 
  guides(fill = FALSE) 
```


---

## Scoring sentiments

.midi[
```{r}
tswift_lyrics %>%
  anti_join(stop_words) %>%
  left_join(get_sentiments("afinn"))
```
]

---

## Assigning a sentiment score

.midi[
```{r}
tswift_lyrics %>%
  anti_join(stop_words) %>%
  left_join(get_sentiments("afinn")) %>%
  filter(!is.na(value)) %>%
  group_by(track_title) %>%
  summarise(total_sentiment = sum(value)) %>%
  arrange(total_sentiment)
```
]
---

## Visualizing sentiment scores

```{r echo  = F, fig.height = 2.5}
tswift_lyrics %>%
  anti_join(stop_words) %>%
  left_join(get_sentiments("afinn")) %>%
  filter(!is.na(value)) %>%
  group_by(track_title) %>%
  summarise(total_sentiment = sum(value)) %>%
  ungroup() %>%
  arrange(total_sentiment) %>%
  mutate(
    total_sentiment_sign = if_else(total_sentiment < 0, "negative", "positive")
  ) %>%
  ggplot(aes(x = reorder(track_title, total_sentiment), y = total_sentiment, fill = total_sentiment_sign)) +
  geom_col() +
  guides(fill = FALSE) +
  coord_flip() +
  labs(x = "", y = "", 
    title = "Total sentiment score of 'Lover' tracks",
    subtitle = "Scored with AFINN sentiment lexicon")
```

---

## ...the code


.small[
```{r eval = F}
tswift_lyrics %>%
  anti_join(stop_words) %>%
  left_join(get_sentiments("afinn")) %>%
  filter(!is.na(value)) %>%
  group_by(track_title) %>%
  summarise(total_sentiment = sum(value)) %>%
  ungroup() %>%
  arrange(total_sentiment) %>%
  mutate(
    total_sentiment_sign = if_else(total_sentiment < 0, "negative", "positive")
  ) %>%
  ggplot(aes(x = reorder(track_title, total_sentiment), y = total_sentiment, 
             fill = total_sentiment_sign)) +
  geom_col() +
  guides(fill = FALSE) +
  coord_flip() +
  labs(x = "", y = "", 
    title = "Total sentiment score of 'Lover' tracks",
    subtitle = "Scored with AFINN sentiment lexicon")
```
]

---

## Additional resources

[Text Mining with R](https://www.tidytextmining.com/)
- Chapter 1: The tidy text format
- Chapter 2: Sentiment analysis with tidy data