---
title: "AE 19: Paris paintings"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
---
## Announcements 

### Tea with a TA 

Hang out with the TAs from STA 199! This is a casual conversation and a fun opportunity to meet the members of the STA 199 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- [**Salvador Arellano**](https://www.linkedin.com/in/salvador-chavero-arellano/), Wed Oct 21, 12p - 1p
  - [Click here](https://forms.gle/CVNc83EsqeuWLj5XA) to sign up
  
- **Morris Greenberg**, Fri Oct 23, 1p - 2p
  - [Click here](https://forms.gle/PgVeB34UhpbvEbqn7) to sign up
  
## Big data and public policy event on Oct 26

## Mid-semester feedback + team work 

## Questions from the video? 

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
```

We will be looking at the Paris Paintings data set in today's application exercise. We'll primarily focus on the variables: 

- `Height_in`: Height (in inches)
- `Width_in`: Width (in inches)
- `landsAll`: If any type of landscape is mentioned (either lands_sc, lands_figs, or lands_ment)

[Click here](https://sta199-fa20-002.netlify.app/slides/paris_codebook.html) for the complete codebook.

```{r load-data, message = FALSE}
paintings <- read_csv("data/paris_paintings.csv")
```

### Exercise 1

Create a scatterplot to visualize the relationship between Height and Width. Color the points based on the whether the painting has any landscape elements. *Note: Be sure to make `landsAll` a factor, so R treats it as categorical variable.*

### Exercise 2

Based on your scatterplot, does the relationship between Height and Width differ between paintings with landscape elements and those without? Briefly explain.  

### Exercise 3

Fit a main effects model using width and whether the painting has landscape elements to predict the height. 

### Exercise 4

Using your model from the previous exercise, 

- Interpret the intercept. 
- Interpret the coefficient (slope) of `Width_in`. 
- Interpret the coefficient (slope) of `landsAll`.

### Exercise 5

Write the equation of the model for 

- paintings with no landscape elements. 
- paintings with landscape elements. 

How do the slopes compare between the two models? How do the intercepts compare? 

### Exericse 6

Now let's consider an interaction term. Fit a linear model using width, whether the painting has landscape elements, and the interaction between the two variables to predict the height. 

### Exercise 7

Write the equation of the model for 

- paintings with no landscape elements. 
- paintings with landscape elements. 

How do the slopes compare between the two models? How do the intercepts compare? 

### Exercise 8

Now let's see how well each model fits the data. Use the `glance` function to calculate $R^2$ and Adj. $R^2$ for the model fit in Ex. 3 and the one fit in Ex. 6. 

### Exercise 9

Based on your answer to the previous exercise, which model is a better fit for the data? Briefly explain.

### Exercise 10

Interpret $R^2$ for the model chosen in the previous exercise.