---
title: "AE 05: Data Wrangling"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
---

## Announcements

- Lab 02 due at 11:59p


## Questions? 


## The way data is displayed matters


## Clone a repo + start a new project

Go to the `ae-05-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. See the [Lab 01](https://sta199-fa20-002.netlify.app/labs/lab-01-hello-r.html#getting-started) for more detailed instructions about cloning a repo and starting a new project. 

### Configure git

Run the following code to configure Git. Fill in your GitHub username and the email address associated with your GitHub account. 

```{r eval=FALSE}
library(usethis)
use_git_config(user.name= "your github username", user.email="your email")
```

## Practice with data wrangling

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

The data set contains all North Carolina bike crash data from 2007-2014. The data was originally downloaded on September 6, 2018 and is available in the dsbox R package. 

Use the code below to load the data from the .csv file

```{r load-data, warning = FALSE, message = FALSE}
ncbikecrash <- read_csv("data/ncbikecrash.csv")
```

### Part 1

**Display the first 10 rows of the ncbikecrash data set.**

```{r}

```


### Part 2 

**Filter for bike crashes in Durham County where the driver age group is 0 to 19.**

```{r}

```

### Part 3

**On which day of the week did the most bike crashes occur? Use the `count` function to get the number of bike crashes for each day (`crash_day`). Display the results in descending order.**

```{r}

```

### Part 4

**Calculate the average estimated driver speed (`driver_est_speed`) based on whether the crash was a hit and run (`hit_run`).**

```{r}

```