---
title: "AE 22: Text analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
---

```{r echo = F}
knitr::opts_chunk$set(warning = F, message = F)
```

```{r echo  = F}
library(tidyverse)
```

## Announcements 

### Assignments 

- [HW 03](https://sta199-fa20-002.netlify.app/hw/hw-03.html) due Nov 4 at 11:59p
- Tomorrow's lab: Project peer review
  - Push a draft of your report to the Github repo by today at 11:59p
  
### Tea with a TA 

Hang out with the TAs from STA 199! This is a casual conversation and a fun opportunity to meet the members of the STA 199 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

[**Caroline Levenson**](https://www.linkedin.com/in/carolinelevenson/), Mon, Nov 2, 1p - 2p

  - [Click here](https://forms.gle/FucAXE6bLeJVyqRY9) to sign up
 
### Submit your questions about statistics and the US election

What questions do you have about statistics and the US election? [Click here](https://forms.gle/6239XcARuuWe4MBw5) to submit your questions **by Friday, Oct 30.** We will discuss some of the questions during the Nov 2 live lecture. 

And...if you're eligible, VOTE!

### Other events 

- [Datathon](https://dukeml.org/datathon/) Oct 31  - Nov 1

### Stats in Spring 2021

- [STA 210: Regression Analysis](https://stat.duke.edu/courses/regression-analysis-0)
- [STA 240: Probability for Statistical Inference, Modeling, and Data Analysis](https://stat.duke.edu/courses/probability-statistical-inference-modeling-and-data-analysis).
  - Pre-req: Mathematics 22, 112L, 122, 122L, 202D, 212, 222


### Project - Draft due Oct 28 at 11:59p

- Write the draft in the `written-report.Rmd` file in the project repo. 

- Draft should include 
   - the research context and motivation
   - exploratory data analysis
   - any inference, modeling, or conclusions.

## Exercises 

For today's AE, we are analyzing the tweets from the statistics experience!

```{r}
# load packages
library(tidyverse)
library(tidytext)
library(stringr)
#library(wordcloud) #word cloud
```

```{r, eval = F}
tweets <- read_csv("data/stats-tweets.csv") %>%
  mutate(tweet_num = 1:nrow())
```


### Exercise 1

Are these Tweets or Tweet threads? Let's see how long the statistics experience tweets are! 

### Exercise 2

Let's use the `unnest_tokens` to make a tidy data frame of words from the tweets. 


### Exercise 3

What are the most common words used in the tweets? Are the most common words interesting? If not, what can we do to make this more interesting? 

### Exercise 4

Make a graph to visualize the top 10 words from your tweets. 

### Exercise 5

Now let's see the general sentiment of the tweets. Use the bing lexicon to identify the top 10 most common words and their sentiments. 

### Exercise 6

Let's visualize the top 10 most common words for each sentiment. Make a plot to display the top 10 most commonly used positive words and negative words. 

### Exercise 7

We can also visualize the most common words using a word cloud. 

```{r}
#library(wordcloud)
#set.seed(10282020)
# _____ %>%
#  anti_join(stop_words) %>%
 # count(word) %>%
#  with(wordcloud(word, n, max.words = 100))
```

### Exercise 8

Based on this analysis, what are some general conclusions you draw about the class's first statistics experiences? 


<hr>

### More with text analysis!

- Text Mining with R: https://www.tidytextmining.com/
- gutenbergr R package: https://cran.r-project.org/web/packages/gutenbergr/vignettes/intro.html
- genius R package: https://github.com/JosiahParry/genius
  - *Note: Limitations in the Docker containers*
- dundlermifflin R package: https://github.com/tbradley1013/dundermifflin
- friends R package: https://emilhvitfeldt.github.io/friends/

