---
title: "AE 20: Number of trail users"
subtitle: "MLR: Inference + conditions"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable
---
## Announcements 

### Tea with a TA 

Hang out with the TAs from STA 199! This is a casual conversation and a fun opportunity to meet the members of the STA 199 teaching team. The only rule is these can't turn into office hours! 

Tea with a TA counts as a statistics experience.

- [**Salvador Arellano**](https://www.linkedin.com/in/salvador-chavero-arellano/), TODAY, 12p - 1p
  - [Click here](https://forms.gle/CVNc83EsqeuWLj5XA) to sign up
  
- **Morris Greenberg**, Fri Oct 23, 1p - 2p
  - [Click here](https://forms.gle/PgVeB34UhpbvEbqn7) to sign up
  
## Big data and public policy event on Oct 26

### Project - Draft due Oct 28

- Write the draft in the `written-report.Rmd` file in the project repo. 

- Draft should include 
   - the research context and motivation
   - exploratory data analysis
   - any inference, modeling, or conclusions.


## Exam 02 

- Released Thursday, Oct 22 at 9a
- Due Sunday, Oct 25 at 11:59p
- Covers statistical inference weeks 05 - 08
- No office hours + Piazza will be inactive during the exam
- If you have questions, email Prof. Tackett (maria.tackett@duke.edu)

## Exercises 

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)

```

The Pioneer Valley Planning Commission collected data in Florence, MA for 90 days from April 5 to November 15, 2005 using a laser sensor, with breaks in the laser beam recording when a rail-trail user passed the data collection station.

- `hightemp`: daily high temperature (in degrees Fahrenheit)
- `volume`: estimated number of trail users that day (number of breaks recorded)
- `fall`: 1: season is Fall, 0: season is not Fall

```{r}
library(mosaicData)
data(RailTrail)
```

### Exercise 1

Fit a linear model using the daily high temperature to predict the estimated number of trail users. Calculate the $R^2$ and $Adj. R^2$ for the model. 

### Exercise 2

Fit a linear model using the daily high temperature and whether it is the fall season to predict the estimated number of trail users. Calculate the $R^2$ and $Adj. R^2$ for the model. 

### Exercise 3

Which model has the higher $R^2$? Which model has the higher $Adj. R^2$? Which model is a better fit for the data? 


### Exercise 4

Let's use the model from Exercise 1. Display the output for the model from Exercise 1 with the 90% confidence interval for the model coefficients.

Interpret the coefficient of `hightemp` in the context of the data. 

### Exercise 5

Interpret the 90% confidence interval for `hightemp` in the context of the data. 

### Exercise 6

Does the intercept have a reasonable interpretation? If so, interpret the intercept in the context of the data. Otherwise, explain why not. 

### Exercise 7

Conduct a hypothesis test to assess if there is a linear relationship the high temperature and number of trail users. State the null and alternative hypotheses, the test statistic, p-value, and conclusion in the context of the data.

### Exercise 8

Is the confidence interval consistent with the conclusion from the hypothesis test in the previous exercise? Briefly explain why or why not. 

### Exercise 9

Let's check the model conditions. Create the diagnostic plots use them to determine if the four model conditions are met.
