---
title: "Strings"
author: "Prof. Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 3, fig.width = 5, dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center") 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

class: middle center

## [Click for PDF of slides](22-strings.pdf)

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

---

```{r echo = FALSE}
states <- c("alabama", "alaska", "arizona", "arkansas", "california", 
            "colorado", "connecticut", "delaware", "florida", "georgia", 
            "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", 
            "kentucky", "louisiana", "maine", "maryland", "massachusetts", 
            "michigan", "minnesota", "mississippi", "missouri", "montana", 
            "nebraska", "nevada", "new hampshire", "new jersey", 
            "new mexico", "new york", "north carolina", "north dakota", "ohio", 
            "oklahoma", "oregon", "pennsylvania", "rhode island",
            "south carolina", "south dakota", "tennessee", "texas", "utah", 
            "vermont", "virginia", "washington", "west virginia", "wisconsin",
            "wyoming")
```


class: middle, center

## `stringr`

---

## `stringr`

In addition to the `tidyverse`, we will use the package `stringr`.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(stringr)
```

`stringr` provides tools to work with character strings.

- Functions in `stringr` have consistent and memorable names

- All begin with `str_` (`str_count`, `str_detect`, `str_trim`, etc)

- All take a vector of strings as their first argument


---


## Preliminaries

Character strings in R are defined by double quotation marks. 

They can include letters, numbers, punctuation, whitespace, etc.

```{r}
string1 <- "STA 199 is my favorite class."
string1
```

You can combine character strings in a vector.

```{r}
string2 <- c("STA 199", "Data Science", "Duke")
string2
```

---

## Include a quotation in a string?

Why doesn't the code below work?

.midi[
```{r eval = FALSE}
string3 <- "I said "Hello" to my class"
```
]

--

To include a double quote in a string, *escape it* using a backslash `\`.

--

.midi[
```{r}
string4 <- "I said \"Hello\" to my class."
```
]

--

What if you want to include an actual backslash?

--

.midi[
```{r}
string5 <- "\\"
```
]

This may seem tedious but it will come up later!

---

## `writeLines`

`writeLines` shows the contents of the string not
including escapes.

.pull-left[
```{r}
string4
writeLines(string4)
```
]
.pull-right[
```{r}
string5
writeLines(string5)
```
]


---

## U.S. States

To demonstrate functions from `stringr` we will use a vector of all 50 states.

.midi[
```{r}
states
```
]

---

## `str_length`

Given a string, return the number of characters.

.midi[
```{r}
string1 <- "STA 199 is my favorite class."
str_length(string1)
```
]

Given a vector of strings, return the number of characters in each string.

.midi[
```{r}
str_length(states)
```
]

--

.pull-left[
- Alabama: 7
- Alaska: 6
- Arizona: 7
- Arkansas: 8
]
.pull-right[
- California: 10
- Colorado: 8
- Connecticut: 11
- ...
]
 

---

## `str_c`

Combine two or more strings.

```{r}
str_c("STA 199", "is", "my", "favorite", "class")
```


--

Use `sep` to specify how the strings are separated.

```{r}
str_c("STA 199", "is", "my", "favorite", "class", sep = " ")
```


---

## `str_to_lower` and `str_to_upper`

Convert the case of a string from lower to upper or upper to lower.

.midi[
```{r}
str_to_upper(states)
```
]


---

## `str_sub`

Extract parts of a string from `start` to `end`, inclusive.

.midi[
```{r}
str_sub(states, 1, 4)
```
]

--

.midi[
```{r}
str_sub(states, -4, -1)
```
]

---

## `str_sub` and `str_to_upper`


Can combine `str_sub` and `str_to_upper` to capitalize each state.

.midi[
```{r}
str_sub(states, 1, 1) <- str_to_upper(str_sub(states, 1, 1))
states
```
]

---

## `str_sort`

Sort a string. Here we sort in decreasing alphabetical order.

.midi[
```{r}
str_sort(states, decreasing = TRUE)
```
]

---

## Regular Expressions

A .vocab[regular expression] is a sequence of characters that allows you to 
describe string patterns. We use them to search for patterns. 

- extract a phone number from text data
- determine if an email address is valid
- determine if a password has the required number of letters, characters, and symbols
- count the number of times "statistics" occurs in a corpus of text
- ...

---

## Regular Expressions

To demonstrate will will use a vector of all of the states bordering North 
Carolina.

```{r}
nc_states <- c("North Carolina", "South Carolina", 
               "Virginia", "Tennessee", "Georgia")
nc_states
```

---

## Basic Match

We can match exactly. 


```{r}
str_view_all(nc_states, "in")
```


---

## Basic Match

Match any character using `.`


```{r}
str_view_all(nc_states, "i.")
```

---

## Anchors

Match the start of a string using `^`

```{r}
str_view_all(nc_states, "^G")
```

---

## Anchors

Match the end of a string using `$`


```{r}
str_view_all(nc_states, "a$")
```

---

## `str_detect`

Determine if a character vector matches a pattern.

```{r}
nc_states
str_detect(nc_states, "a")
```


---

## `str_subset`

Select elements that match a pattern.

```{r}
str_subset(nc_states, "e$")
```


---

## `str_count`

How many matches are there in a string?


```{r}
nc_states
str_count(nc_states, "a")
```


---

## `str_replace`

Replace first match with new strings.


```{r}
str_replace(nc_states, "a", "-")
```


---

## `str_replace_all`

Replace all matches with new strings.

```{r}
str_replace_all(nc_states, "a", "-")
```



---

## Many Matches

The regular expressions below match more than one character.

- Match any digit using `\d` or `[[:digit:]]`
- Match any whitespace using `\s` or `[[:space:]]`
- Match f, g, or h using `[fgh]` 
- Match anything but f, g, or h using `[^fgh]`
- Match lower-case letters using `[a-z]` or `[[:lower:]]`
- Match upper-case letters using `[A-Z]` or `[[:upper:]]`
- Match alphabetic characters using `[A-z]` or `[[:alpha:]]`

Remember these are regular expressions! To match digits you'll need to *escape*
the string, so use `"\\d"`, not `"\d"`

---

## Additional resources

- `stringr` website: https://stringr.tidyverse.org/
- `stringr` [cheat sheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/strings.pdf)
- Regular Expressions [Cheat Sheet](https://rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf)
- [Chapter 14: Strings](https://r4ds.had.co.nz/strings.html#matching-patterns-with-regular-expressions) in R for Data Science
