---
title: "Lab 08: Logistic regression"
subtitle: "due Wed, Nov 11 at 11:59p"
output: 
  tufte::tufte_html:
    highlight: pygments
    css: "sta199-labs.css"
link-citations: yes
---

```{r include=FALSE}
library(tidyverse)
library(tufte)
library(knitr)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  digits = 2
  )
knitr::opts_chunk$set(eval = FALSE)
```

# Intensive Care Unit

The Intensive Care Unit (ICU) is the specialized unit of a hospital that focuses
on extremely severe and life-threatening medical issues (septic shock, organ 
failure, major trauma, etc). Patients admitted to the ICU require constant 
care and monitoring by highly-trained medical professionals and the use of 
advanced equipment.

Researchers are interested in the survival of patients admitted to the ICU. 
Patients with a low predicted probability of survival can be triaged to more 
medical resources and care, and patients with a high predicted probability of 
survival can be triaged to another unit of the hospital. 

The file `icu.csv` contains information for a random sample of 150 patients at 
an Intensive Care Unit. Researchers are interested in predicting patient 
survival. This data is from the [Data and Story Library (DASL)](http://lib.stat.cmu.edu/DASL/Datafiles/ICU.html).

- `id`: patient ID code
- `survive`: survival status (0 = patient died, 1 = patient survived to 
discharge)
- `age`: age (years)
- `sex`: sex (0 = male, 1 = female)
- `infection`: infection status (0 = no infection, 1 = infection suspected)
- `sysbp`: systolic blood pressure (mm of Hg)
- `pulse`: heart rate (beats per minute)
- `emergency`: admission status (0 = elective admission, 1 = emergency 
admission)

# Exercises

1. Create three effective, well-labeled visualizations to explore the 
relationship between the variables below. Provide a brief one or two sentence
comment on each.

    - `age` and `survive`
    - `emergency` and `survive`
    - `sysbp` and `survive`
  
2. Fit a logistic regression model with `survive` as the response variable with the predictors `age`, 
`sysbp` and `emergency`. Report the model output in tidy format.

3. Write out the equation of the fitted logistic regression model. You can use "pred-log-odds" for the response variable.

4. Provide an interpretation of $\hat{\beta}_{\text{emergency}}$ in the context
of the problem. Your interpretation should involve **odds** or **odds ratios**,
not **log odds**.

5. Provide an interpretation of $\hat{\beta}_{\text{age}}$ in the context of the
problem. Your interpretation should be in terms of **odds** or **odds ratios**,
not log odds.

6. An 87 year old patient is admitted in an emergency situation with a systolic
blood pressure of 88 mm of Hg. What is this patient's predicted log-odds of
survival?

7. What is this patient's predicted probability of survival?

8. The patient described in Exercises #6 and #7 survived. Is this result 
surprising? Answer with justification.

9. A patient is admitted in an emergency situation with a systolic blood 
pressure of 80 mm of Hg. How young must the patient be in order to have
a predicted probability of survival at least 0.50 based on the logistic
regression model?

10. The file `icu_test.csv` contains a test dataset of 50 patients not included
in the original data. Fill in the code below to find the model predictions for the 50 patients, and then calculate their predicted probabilities. 


```{r eval =  F}
icu_test <- read_csv("data/icu_test.csv")
_____ <- augment(_____, newdata = icu_test)
```

# Submission

Knit to PDF to create a PDF document. Knit and commit all remaining changes, 
and push your work to GitHub. Make sure all files are updated on your GitHub 
repo.

Please only upload your PDF document to Gradescope. Associate the "Overall"
graded section with the first page of your PDF, and mark where each answer is 
to the exercises. If any answer spans multiple pages, then mark all pages.