---
title: "Lab 02 - Data wrangling and visualization"
subtitle: "Exploring college majors"
date: "due Thursday, Jan 30 at 11:59p"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: "./sta199-labs.css"
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

```{marginfigure}
If you are curious about how raw data from the ACS were cleaned and prepared, 
see [**the code**](https://github.com/fivethirtyeight/data/blob/master/college-majors/college-majors-rscript.R) that the FiveThirtyEight authors used 
(be warned: it's a bit outside of the scope of this course!).
```

Often times, the first step in using statistics to turn information into 
knowledge is to summarize and describe the raw information - the data. In this 
lab we explore data on college majors and earnings, specifically the data behind 
the FiveThirtyEight story [**The Economic Guide To Picking A College Major**](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/). 

These data originally come from the American Community Survey (ACS) 2010-2012 
Public Use Microdata Series. 
There are many considerations that go into picking a major; two of them are
earnings potential and employment prospects. They are important, but they don't 
tell the whole story. Keep this in mind as you analyze the data.

# Getting Started

## Clone Assignment Repo

- First, go to the link for the assignment in order to create your personal 
  private repository corresponding to the assignment. This link may be found at
  [**https://classroom.github.com/a/27agHlzr**](https://classroom.github.com/a/27agHlzr). 
  Accepting the assignment creates a repository similar to `lab02-[github_name]`.
- Next, navigate to the course GitHub website, and click on the repository 
  corresponding to your account. For example, mine 
  would be `lab02-shawnsanto`. Or, simply go to the repository you just created 
  (there is a link provided when you created the repository).
- Click on the green `Clone or download` button in that repository, and copy 
  the git URL (it should end in .git)
- Go to RStudio Cloud and into **the STA 199 course workspace**. Create a 
  **New Project from Git Repo** and copy the git URL from your personal 
  workspace. Make sure "Add packages from the base project" is checked.
- Click OK, and you should see the contents from your GitHub repo in the `Files`
  pane in RStudio. 

## Configure git

Remember that we first have to configure RStudio Cloud to talk to GitHub. An 
easy way to do this is to use the `use_git_config()` function from the 
`usethis` package. 
Type the following lines of code in the **`Console`** in RStudio filling in your 
name and email address.

```{r eval=FALSE}
library(usethis)
use_git_config(user.name="Your name", user.email="your-git-hub-email@email.edu")
```

## Project name: 

Currently your project is called `Untitled Project`. Update the name of your 
project to be "Lab 02 - Data wrangling and visualization".

## Packages

In this lab we will work with the `tidyverse` and `fivethirtyeight` packages. 
Load the packages into the **`Console`** (they have already been installed 
for you).

```{r}
library(tidyverse)
library(fivethirtyeight)
```

# Warm up

Before we introduce the data, let's do a few things, which hopefully 
should be review.

## YAML: 

Open the R Markdown (Rmd) file in your project, change the author name to your 
name, and knit the document.

## Stage, commit, and push changes:

- Go to the **Git** pane in RStudio. 
- View the **Diff** and confirm that you are happy with the changes.
- **Stage** the files you want to commit.
- Add commit message "Updated name" in the **Commit message** box and hit **Commit**.
- Click on **Push**. This will prompt a dialog box where you first need to 
  enter your GitHub user name and password.
For the remainder of this lab, we won't tell you when to commit -- it is up to 
you to commit at appropriate intervals with meaningful commit messages. Be sure 
to commit **at least three times** during this lab.

# Load the data

The data frame we will be working with today is called `college_recent_grads` 
and it's in the `fivethirtyeight` package.
To find out more about the dataset, type the following in your `Console`: 
`?college_recent_grads`. A question mark before the name of an object will 
bring up its help file. This command must be run in the `Console`.
`college_recent_grads` is a tidy **data frame**, with each row representing 
an **observation** and each column representing a **variable**.
Take a quick peek at your data frame and view its dimensions with 
the `glimpse()` function.

```{r glimpse}
glimpse(college_recent_grads)
```

The description of the variables, i.e. the **codebook**, is given below.

| Header                        |  Description
|:----------------|:--------------------------------
|`rank`                         | Rank by median earnings
|`major_code`                   | Major code, FO1DP in ACS PUMS
|`major`                        | Major description
|`major_category`               | Category of major from Carnevale et al
|`total`                        | Total number of people with major
|`sample_size`                  | Sample size (unweighted) of full-time, year-round ONLY (used for earnings)
|`men`                          | Male graduates
|`women`                        | Female graduates
|`sharewomen`                   | Women as share of total
|`employed`                     | Number employed (ESR == 1 or 2)
|`employed_full_time`           | Employed 35 hours or more
|`employed_part_time`           | Employed less than 35 hours
|`employed_full_time_yearround` | Employed at least 50 weeks (WKW == 1) and at least 35 hours (WKHP >= 35)
|`unemployed`                   | Number unemployed (ESR == 3)
|`unemployment_rate`            | Unemployed / (Unemployed + Employed)
|`median`                       | Median earnings of full-time, year-round workers
|`p25th`                        | 25th percentile of earnings
|`p75th`                        | 75th percentile of earnings
|`college_jobs`                 | Number with job requiring a college degree
|`non_college_jobs`             | Number with job not requiring a college degree
|`low_wage_jobs`                | Number in low-wage service jobs

The `college_recent_grads` data frame is a trove of information. 

Let's think about some questions we might want to answer with these data.
- Which major has the lowest unemployment rate?
- Which major has the highest percentage of women?
- How do the distributions of median income compare across major categories?
- Do women tend to choose majors with lower or higher earnings?

In the next section we will answer these questions.

# Data wrangling and visualization

## Which major has the lowest unemployment rate?

In order to answer this question all we need to do is sort the data. We can use 
the `arrange()` function to do this, and sort it by the `unemployment_rate` 
variable. By default, `arrange()` sorts in ascending order, which is what we 
want here -- we're interested in the major with the *lowest* unemployment rate.

```{r lowest-unemp}
college_recent_grads %>%
  arrange(unemployment_rate)
```

This gives us what we wanted, but not in an ideal form. First, the name of the 
major barely fits on the page. Second, some of the variables are not that 
useful (e.g. `major_code`, `major_category`) and some we might want front 
and center are not easily viewed (e.g. `unemployment_rate`).
We can use the `select()` function to choose which variables to display and 
in which order:

```{marginfigure}
Note how easily we expanded our code with adding another step to our pipeline, 
with the pipe operator: `%>%`.
```

```{r lowest-unemp-select}
college_recent_grads %>%
  arrange(unemployment_rate) %>%
  select(rank, major, unemployment_rate)
```

This is looking better, but do we really need all those decimal places in the 
unemployment variable? Not really!

There are two ways we can address this problem. One would be to round the 
`unemployment_rate` variable in the dataset or we can change the number of 
digits displayed, without touching the input data. Below are instructions for 
how you would do both of these:

- **Round the variable itself**: We create a new variable with the `mutate()` 
  function. In this case, we're overwriting the existing `unemployment_rate` 
  variable, by `round`ing it to `4` decimal places.
  
```{r}
college_recent_grads %>%
  arrange(unemployment_rate) %>%
  select(rank, major, unemployment_rate) %>%
  mutate(unemployment_rate = round(unemployment_rate, digits = 4))
```

- **Change the displayed number of digits without touching data**: We can add 
  an option to our R Markdown document to change the displayed number of digits 
  in the output. To do so, add a new chunk, and set:
  
```{r}
options(digits = 2)
```

Note that the `digits` in `options` is scientific digits, and in `round` they 
are decimal places. If you're thinking "Wouldn't it be nice if they were 
consistent?"...so do we. You don't need to do both of these, that would be 
redundant. The next exercise asks you to choose one.

1. Which of these options, changing the input data or altering the number of 
   digits displayed without touching the input data, is the better option? 
   Explain your reasoning. Implement the option you chose.
   
## Which major has the highest percentage of women?

To answer such a question we need to arrange the data in descending order. 
For example, if earlier we were interested in the major with the highest 
unemployment rate, we would use the following:

```{marginfigure}
The `desc()` function specifies that we want `unemployment_rate` in descending 
order.
```

```{r}
college_recent_grads %>%
  arrange(desc(unemployment_rate)) %>%
  select(rank, major, unemployment_rate)
```

2. Using what you've learned so far, arrange the data in descending order with 
   respect to proportion of women in a major, and display only the major, the 
   total number of people with that major, and proportion of women. 
   Show only the top 3 majors by adding `slice(1:3)` at the end of the pipeline.
   
## How do the distributions of median income compare across major categories?

```{marginfigure}
A percentile is a measure used in statistics indicating the value below which 
a given percentage of observations in a group of observations fall. 
For example, the 20th percentile is the value below which 20% of the observations 
may be found. (Source: [Wikipedia](https://en.wikipedia.org/wiki/Percentile)
```

There are three types of incomes reported in this data frame: `p25th`, `median`,
and `p75th`. These correspond to the 25th, 50th, and 75th percentiles of the 
income distribution of sampled individuals for a given major.

3. Why do we often choose the median, rather than the mean, to describe the 
   typical income of a group of people?
   
In order to answer our question about the distributions of median income across 
major categories, we need to group the data by `major_category`. Then, we need 
a way to summarize the distributions of median income within these groups. 
This decision will depend on the shapes of these distributions. Let's create
a visualization to see the shapes of these distributions.

Function `ggplot()` will allow us to create our plots. The first argument is 
the data frame, and the next argument gives the mapping of the variables of 
the data to the `aes`thetic elements of the plot.

Let's start simple and take a look at the distribution of all median incomes, 
without considering major categories.

```{r}
ggplot(data = college_recent_grads, mapping = aes(x = median)) +
  geom_histogram()
```
Along with the plot, we get a message:

```
`stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
```

This is telling us that we might want to reconsider the binwidth we chose for 
our histogram -- or more accurately, the binwidth we didn't specify. It's good 
practice to always think in the context of the data and try out a few binwidths
before settling on your final choice. You might ask yourself: "What would be a 
meaningful difference in median incomes?" \$1 is obviously too little, 
\$10000 might be too high.

4. Plot two histograms of median incomes, one with a bin width of \$1000 and 
one with a bin width of \$5000. Select one and explain your reasoning for your 
choice. Argument binwidth is for the `geom_histogram()` function. So to specify 
a binwidth of \$1000, you would use `geom_histogram(binwidth = 1000)`.
We can also calculate summary statistics for this distribution using the 
`summarise()` function:

```{r}
college_recent_grads %>%
  summarise(min  = min(median), 
            max  = max(median),
            mean = mean(median), 
            med  = median(median),
            sd   = sd(median), 
            q1   = quantile(median, probs = 0.25),
            q3   = quantile(median, probs = 0.75))
```

5. Based on the shape of the histogram you created in the previous exercise, 
   determine which of these summary statistics is useful for describing the 
   distribution. Write up your description (remember shape, center, spread, 
   any unusual observations) and include the summary statistic output as well.

Next, we facet the plot by major category.

```{r}
ggplot(data = college_recent_grads, mapping = aes(x = median)) +
  geom_histogram() +
  facet_wrap( ~ major_category, ncol = 4)
```

6. Plot the distribution of `median` income using a histogram, faceted by 
   `major_category`. Use the `binwidth` you chose in the earlier exercise.
Now that we've seen the shapes of the distributions of median incomes for each 
major category, we should have a better idea for which summary statistic to use 
to quantify the typical median income.

7. Which major category has the highest typical (you'll need to decide what 
   this means) median income? Use the partial code below, filling it in with 
   the appropriate statistic and function. Also note that we are looking for 
   the highest statistic, so make sure to arrange your resulting data frame in 
   the correct order. 
   
```{r eval=FALSE}
college_recent_grads %>%
  group_by(major_category) %>%
  summarise(___ = ___(median)) %>%
  arrange(___)
```

8. Which major category is the least popular in this sample? To answer this 
   question we use a new function called `count()`, which first groups the data 
   and then counts the number of observations in each category (see below). 
   Arrange the results so that the major with the lowest observations is 
   in row 1 of the resulting data frame.
   
```{r}
college_recent_grads %>%
  count(major_category)
```

## All STEM fields aren't the same

One section of the [FiveThirtyEight story](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/) 
is titled "All STEM fields aren't the same". Let's see if this is true.
First, let's create a new vector called `stem_categories` that lists the major 
categories that are considered STEM fields. Ask your TA what function `c()` does
if you are not sure. Also, check the help with `?c`.

```{r}
stem_categories <- c("Biology & Life Science",
                     "Computers & Mathematics",
                     "Engineering",
                     "Physical Sciences")
```

We can use this to create a new variable in our data frame indicating whether a
major is STEM or not.

```{r}
college_recent_grads <- college_recent_grads %>%
  mutate(major_type = ifelse(major_category %in% stem_categories, "stem", "not stem"))
```

Let's unpack this: with `mutate()` we create a new variable called 
`major_type`, which is defined as `"stem"` if the `major_category` is in the 
vector called `stem_categories` we created earlier, and as `"not stem"` 
otherwise.
`%in%` is a **logical operator**. Other logical operators that are commonly 
used are

| Operator                   | Operation
|:----------------|:--------------
|`x < y`                     | less than
| `x > y`                    | greater than
| `x <= y`                   | less than or equal to
| `x >= y`                   | greater than or equal to
| `x != y`                   | not equal to
| `x == y`                   | equal to
| `x %in% y`                 | group membership
| `x | y`                    | or
| `x & y`                    | and
| `!x`                       | not

Here is a small example showing how the operator `%in%` works:

```{r eval=TRUE}
x <- c("red", "white", "blue")
y <- c("green", "red", "red", "orange", "white", "black", "purple", "blue")
y %in% x
```

We can use logical operators to `filter()` our data. Here will filter based on
the conditions that the major type is STEM and the median salary is less than
\$36,000.

```{r}
college_recent_grads %>%
  filter(
    major_type == "stem",
    median < 36000
  )
```

9. Which STEM majors have median salaries equal to or less than the median for 
   all majors' median earnings? Your output should only be a table that shows 
   the major name, median, 25th percentile, and 75th percentile earning for 
   that major, and should be sorted such that the major with the highest median 
   earning is on top. **Do not include narrative; the table itself will suffice**.

## What types of majors do women tend to choose?

10. Create a scatterplot of median income vs. proportion of women in that 
    major. Set the point color as to whether the major is in a STEM field or 
    not. Describe the association between these three variables. 
    Make sure your figure is an appropriate size (make it have square dimensions).